\documentclass{article}
\usepackage{enumerate}
\usepackage[utf8]{inputenc}

\title{14.01 Recitation 7}
\author{Robert Durfee}
\date{November 2017}

\begin{document}

\maketitle

\section*{Problem 1}

\begin{enumerate}
    \item \textbf{Under perfect discrimination, there is no dead weight loss. True or false?}
    
    True. All of the consumer surplus is extracted. Each consumer pays their willingness to pay. Marginal revenue equals the demand.
    
    \item \textbf{Under oligopolistic competition, consumers are weakly worse off than in a perfectly competitive market. True or False?}
    
    True. When there isn't perfect competition, there is market power. The consumer is forced to pay what the seller wants.
    
\end{enumerate}

\section*{Problem 2}

There are $n$ firms where $n \geq 2$. 

$$P(Q)=\alpha-Q$$
$$Q=q_{1}+q_{2}+...+q_{n}$$
$$C(q_{i})=\frac{1}{2}q_{i}^{2}$$

\begin{enumerate}
    \item \textbf{What is firm i's profit?}
    
    $$\pi_{i}(q_{1},...,q_{n})=q_{i}(\alpha-[q_{1}+...+q_{n}])-\frac{1}{2}q_{i}^{2}$$
    
    \item \textbf{Find the Cournot Equilibrium.}
    
    \begin{enumerate}[i]
        \item Assume that firm 1 will know all the quantities produced by the other firms ($q_{1},...,q_{n}$).
        \item Find the best response of firm 1. You have to maximize the profit.
        $$\frac{\partial \pi_{1}(q_{1},...,q_{n})}{\partial q_{1}}=\alpha-[q_{1}+...+q_{n}]-q_{1}-q_{1}=0$$
        $$q_{i}^{*}(q_{1},...,q_{i-1},q_{i+1},...,q_{n})=\frac{\alpha-\sum_{j\neq i} q_{j}}{3}$$
        $$\alpha-3q_{i}-\sum_{j\neq i}q_{j}=0$$
        The cost for each firm is the same.
        \item Solve the equations.
        $$\alpha - 3q^{*}-\sum_{j\neq i}q^{*}=0$$
        $$\alpha - 3q^{*}-(n-1)q^{*}$$
        $$q^{*}=\frac{\alpha}{n+2}$$
    \end{enumerate}
    
    \item \textbf{Find the price.}
    
    $$P(Q^{*})=\alpha-Q^{*}$$
    $$P(Q^{*})=\alpha-n q^{*}$$
    $$P(Q^{*})=\alpha-n \frac{\alpha}{n+2}=\alpha\frac{2}{n+2}$$
    
\end{enumerate}

\section*{Problem 3}

Same as problem 2, but with non-symmetrical cost equation:

$$C(q_{i})=c_{1}q_{i}+\frac{c_{2}q_{i}^{2}}{2}$$

\section*{Problem 4}

There are two firms, A and B.

$$P(Q)=10-Q$$
$$Q=Q_{A}+Q_{B}$$
$$C_{A}(Q_{A})=2Q_{A}$$
$$C_{B}(Q_{B})=4Q_{B}$$

\begin{enumerate}
    \item \textbf{Find the joint profit maximizing price, quantity, and total profit.}
    
    Firm B shuts down.
    
    $$MR=MC_{A}$$
    $$MR=10-2Q_{A}^{*}$$
    $$MC_{A}=4$$
    $$Q^{*}=4$$
    $$P^{*}=6$$
    $$\pi=P^{*}Q^{*}-2Q^{*}=16$$
    
    \item \textbf{Find the equilibrium price.}
    
    \begin{enumerate}[i]
        \item Find the profit for each firm.
        
        $$\pi_{A}=(10-Q_{A}-Q_{B})Q_{A}-2Q_{A}$$
        $$\pi_{B}=(10-Q_{A}-Q_{B})Q_{B}-4Q_{B}$$
        
        \item Find the best response for each firm.
        
        $$\frac{\partial \pi_{A}}{\partial Q_{A}}=8-Q_{B}-2Q_{A}=0$$
        $$\frac{\partial \pi_{B}}{\partial Q_{B}}=6-Q_{A}-2Q_{B}=0$$
        
        \item Solve the system of equations.
        
        $$Q_{A}^{*}=10/3\ Q_{B}^{*}=4/3$$
        $$P^{*}=16/3$$
        
    \end{enumerate}
    
    
    
\end{enumerate}

\end{document}
