\documentclass{article}
\usepackage{tikz}
\usepackage{float}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{indentfirst}
\usepackage{siunitx}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{ {Images/} }
\usepackage{float}
\usepackage{mhchem}
\usepackage{chemfig}
\allowdisplaybreaks

\begin{document}

\begin{enumerate}[1.]
  \item Right away, I noticed that I really did look at my slides quite often in
    my second run of my previous project talk. I did not notice that I was doing
    this while I was speaking, but I can see how it can be distracting. In some
    cases I think it was justified and helped guide the audience to look at the
    slides. In other cases, it was pointless.
  \item I felt that my volume, diction, and vocal variation was good during my
    introduction. This occurred around 00:31. I thought that my gestures looked
    natural during the presentation, especially with the second build slide
    around 01:05. Although in retrospect I wouldn't have looked at my slides as
    much while presenting this slide, I think my hand gestures fit well with the
    content on the screen.
  \item All throughout the presentation, particularly with the final slide
    around 02:18, I interacted a little too much with my slides. I knew my
    slides well, but I often looked at the screen when I was describing them. I
    should've kept eye contact with the audience during this time. Also, I feel
    I could add a little more excitement into the presentation. Particularly
    around 03:21 with the conclusion.
  \item The big thing that I will focus on for my next talk is not looking at my
    slides nearly as much. To do this, I think I will try to study my slide
    progression more before giving the presentation and to position myself in
    such a way that it becomes difficult for me to glance at the slides. I also
    think I could try to add a little more excitement about my topic, but I
    think this will have to come organically. Perhaps better topic choices will
    help with this also.
\end{enumerate}

\end{document}
