\documentclass{article}
\usepackage{tikz}
\usepackage{float}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{indentfirst}
\usepackage{siunitx}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{ {Images/} }
\usepackage{float}
\usepackage{mhchem}
\usepackage{chemfig}
\allowdisplaybreaks

\title{ 5.111 Lecture 31 }
\author{ Robert Durfee }
\date{ December 1, 2017 }

\begin{document}

\maketitle

\section{ Nickel Complexes }

Looking at $\ce{ [NiCl_{6}]^{4-} }$. This absorbs long (red) wavelenght so
$\Delta_{O}$ is small and $\ce{ Cl- }$ is a weak ligand. If you swap this will
$\ce{ H_{2}O }$, ($\ce{ [ Ni( H_{2}O )_{6} ]^{2+} }$) it will absorb a shorter
wavelgnth with a larger $\Delta_{O}$ and thus $\ce{ H_{2}O }$ is a stronger
ligand.

Looking at $\ce{ Ni-( dmgH )_{2} }$, there is a much smaller wavelength absorbed
and thus $\Delta_{O}$ is much larger. However, the complex is square planar,
which causes a larger splitting of the d-orbital energies.

\section{ Tetrahedral Geometry Crystal Field Theory }

There is more repulsion between the ligand point charges and the orbitals that
are $45^{o}$ off the axis ($d_{xy}$, $d_{yz}$, and $d_{zx}$) than the others.

The tetrahedral crystal field splitting energy is smaller than for octahedral
complexes because the point charges are not directed at any orbital set. Because
$\Delta_{T}$ is small, many tetrahedrals are high spin. We can assume they are
all high spin. The ratio between the energies is maintained as before.
(Remember: high spin means don't pair electrons until all orbitals are full).

\section{ Square Planar Geometry Crystal Field Theory }

The $d_{xy}$ and $d_{x^{2}-y^{2}}$ are destabilized the most. Lots of repulsion
because ligand point charges direct at orbitals. The $d_{z^{2}}$ has the least
amount of repulsion. The most unstable is $d_{x^{2}-y^{2}}$ and and $d_{xy}$ is
the second. The rest are about the same at lower repulsion.

\section{ Square Pyramidal Geometry Crystal Field Theory }

$d_{z^{2}}$ is destabilized for square pyramidal. $d_{xz}$ and $d_{yz}$ are
destabilized for square pyramidal. $d_{x^{2}-y^{2}}$ and $d_{xz}$ are not
degenerate for square pyramidal. 

\end{document}

