\documentclass{article}
\usepackage{tikz}
\usepackage{float}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{indentfirst}
\usepackage{siunitx}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{ {Images/} }
\usepackage{float}
\usepackage{mhchem}
\usepackage{chemfig}
\allowdisplaybreaks

\title{ 24.00 Recitation 2 }
\author{ Robert Durfee }
\date{ February 16, 2018 }

\begin{document}

\maketitle

\section{ Design Argument }

Something is very complicated (like an eye). If any part of it were any
different, if would not be able to exist. Thus, there must have been a designer.
Is this a 'God of the gaps' argument? Is it something we can't explain so we
just say, "God must've done that!" Darwin came up with a better explanation, a
scientific explanation.

\section{Fine Tuning Argument}

Everything is so perfect for life to exists in this universe. Someone must've
arranged the constants such that life could exist. But, could it have been a
coincidence? No, this requires an explanation. Science longs to explain
anomalies such as this.

There are some who think that we do not need to explain it at all because it is
the only observation that we could've possibly made. Then there is the firing
squad example. All bullets miss you, you observe this, this is the only possible
explanation, yet it still begs for an explanation.

\section{Multiverse Explanation}

All possible universes exists simultaneously. Why are constants fine tuned?
Because we are only one of many universes, many of which do not permit life.

\section{The Problem of Evil Argument}

If God exists, and he is all good, there is no good reason, but there is evil.
You can reject the fact there is evil, everything is just a level of good. But
this is a fundamental part of theism. To explain this, there must be a good
reason for evil to exist. An explanation for evil is discipline, to teach a
lesson.

\end{document}

