\documentclass{article}
\usepackage{tikz}
\usepackage{float}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{indentfirst}
\usepackage{siunitx}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{ {Images/} }
\usepackage{float}
\usepackage{mhchem}
\usepackage{chemfig}
\allowdisplaybreaks

\title{ 14.01 Recitation 9 }
\author{ Robert Durfee }
\date{ December 1, 2017 }

\begin{document}

\maketitle

\section{ True or False }

\begin{enumerate}[1.]
    \item A firm with production $q = L + 6K$ will always hire either labor or
        capital but never both at the same time.

        \textbf{False.} The hiring of both labor and capital is dependent upon
        the ratio between the prices of labor and capital.
\end{enumerate}

\section{ Social Welfare and Redistribution }

Consider an economy with only one good. There are three people in the economy:
A, B, and C. They all have the same utility:
$$ U = \sqrt{ \rm{ units\ of\ good } } $$

The social welfare function for this society is the sum of the individual
utilities of the three people. A has 400 units, B has 100 units, and C has 16
units of the good. These are the initial endowments of the good for each of the
individuals.

\begin{enumerate}[1.]
    \item If each agent simply consumes his own endowment, what is the social
        welfare?
        $$ SW = \sqrt{ 400 } + \sqrt{ 100 } + \sqrt{ 16 } = 34 $$

    \item Now, assume that the government redistributes the goods by taking 175
        units from A and gives it to B. But there is loss in this transaction of
        79 units. What is the social welfare?
        $$ SW = \sqrt{ 400 - 175 } + \sqrt{ 100 + 175 - 79 } + \sqrt{ 16 } = 33 $$

    \item Now, instead of transfering from A to B, 175 units are transferred
        from A to C with a loss of 91 units.
        $$ SW = \sqrt{ 400 - 175 } + \sqrt{ 100 } + \sqrt{ 16 + 175 - 91 } = 35 $$

    \item Now, the social welfare is the product of the individual utilities.
        \begin{enumerate}[a.]
            \item $SW = 800$
            \item $SW = 1260$
            \item $SW = 1500$
        \end{enumerate}

        In this case, the social welfare increases with all redistribution to
        the poorer agent. In the first case, the redistribution to the poorer
        agent did not always result in an increase in social welfare. This
        social welfare function ignores the cost of transferring between agents.

    \item Now, the social welare is:
        $$ U = U_{A}^{2} + U_{B}^{2} + U_{C}^{2} $$
        \begin{enumerate}[a.]
            \item $SW = 516$
            \item $SW = 437$
            \item $SW = 425$
        \end{enumerate}

        In this case, the social welfare decreases with redistribution because
        of the cost of transfering the wealth. As this cost goes up, the social
        welfare goes down. This social welfare function only cares about the
        loss from transfer, not about the redistribution.
        

\end{enumerate}

\section{ Social Welfare and Redistribution }

Consider an economy that has two types of individuals (rich and poor) and the
only difference between them is their wages. 
$$ w_{r} = 9 $$
$$ w_{p} = 3 $$

Suppose one quarter of the population is rich and the rest are poor. The
individuals have a utility function which considers consumption and leisure:
$$ U( c, l ) = c - \frac{ 2 }{ 3 } \cdot ( 168 - l )^{3/2} $$

And their income is given by:
$$ Y = w_{i} \cdot ( 168 - l ) $$

\begin{enumerate}[1.]
    \item What is the budget constraint for the two types of individuals?

        The budget constraint for each individual will have an x-axis intercept
        equal to 168 hours of leisure and the rich individual will have a y-axis
        intercept equal to $9 \cdot 168$ and the poor individual will have a
        y-axis intercept equal to $3 \cdot 168$ units of consumption.

    \item Find the optimal level of labor, consumption, and leisure.

        Equate marginal rate of substitution of consumption and leisure equal to
        the prices of leisure and consumption to find optimal leisure.
        $$ \frac{ \partial U / \partial l }{ \partial U / \partial c } = w $$

        $$ l^{*} = 168 - w^{2} $$
        $$ c^{*} = w \cdot ( 168 - l^{*} ) = w \cdot w^{2} = w^{3} $$
        $$ h^{*} = 168 - l^{*} $$

        Substitute the values for wage to get the values for each type:
        $$ l^{*}_{p} = 159 $$
        $$ h^{*}_{p} = 9 $$
        $$ c^{*}_{p} = 27 $$
        $$ l^{*}_{r} = 87 $$
        $$ h^{*}_{r} = 81 $$ 
        $$ c^{*}_{r} = 729 $$

\end{enumerate}

\end{document}

